### 2016-07-06 introduction.tex
edits

In this paper, we propose a program logic for weakly-isolated
transactions that realizes this goal.  In particular, we develop a set
of syntax-directed compositional proof rules that allow developers to
build correctness proofs for transactional programs in the presence of
a weakly-isolated concurrency control mechanism.  A key novelty of our
approach is that it is parametric over the \emph{isolation semantics}
of transactions in the program, as well as the \emph{consistency
semantics} of the underlying store. In concrete terms, this means
that, unlike recent work focused on reasoning about programs under
weak consistency~\cite{gotsmanpopl16, redblueatc, ecinec}, our system
model does not assume a minimum or predefined set of consistency or
isolation levels. Instead, our operational semantics admits
declarative specifications of transaction isolation and store
consistency, and generates executions that are guaranteed to satisfy
these specifications. The result is a flexible system model that is
general enough to incorporate the semantics of a range of isolation
levels on a variety of stores (\eg~the sequentially consistent store
of~\cite{adyaphd}, or the causally consistent store
of~\cite{gotsmanpopl16}, etc).  Our key technical contribution is thus
a set of proof rules that demonstrate that a program with a given
selection of isolation levels for its transactions preserves its
invariants when executed on a store equipped with certain consistency
guarantees (\eg~causal consistency).

The paper makes the following contributions:

\begin{itemize}
  \item We develop a semantics for a core language equipped with
    weakly-isolated transactions, demonstrating that a general
    parametric isolation semantics can be expressed as
    well-formedness constraints on a program's execution.
  \item We present a compositional proof system for this language
    capable of relating high-level application invariants to the structure
    of traces induced by the operational semantics, in which transactions
    are associated with specific weak isolation levels.
  \item We define a \emph{maximum visibility principle} to reconcile
    store consistency guarantees with transaction isolation
    requirements, and use it to instantiate our operational model and
    the proof system for various kinds of (weakly-consistent) data
    stores.
  \item Case studies modeled after real-world scenarios demonstrate
    the applicability and utility of our proof methodology.
\end{itemize}

\noindent Our results provide the first (to the best of our knowledge)
mechanism that precisely and uniformly relates high-level program
invariants to low-level weak isolation and weak consistency
guarantees, thereby allowing weakly-isolated transactions to enjoy the
same rigorous reasoning capabilities as their strongly-isolated
(serializable) counterparts.

>>>>

In this paper, we propose a program logic for weakly-isolated
transactions that realizes this goal.  In particular, we develop a set
of syntax-directed compositional proof rules that allow developers to
build correctness proofs for transactional programs in the presence of
a weakly-isolated concurrency control mechanism.  A key novelty of our
approach is that it is parametric over the \emph{isolation semantics}
of transactions in the program, as well as the \emph{consistency
semantics} of the underlying store. In concrete terms, this means
that, unlike recent work focused on reasoning about programs under
weak consistency~\cite{gotsmanpopl16, redblueatc, ecinec}, our system
model does not assume a minimum or predefined set of consistency or
isolation levels. Instead, our operational semantics admits
declarative specifications of transaction isolation and store
consistency, and generates executions that are guaranteed to satisfy
these specifications. The result is a flexible system model that is
general enough to incorporate the semantics of a range of isolation
levels on a variety of stores (\eg~the sequentially consistent store
of~\cite{adyaphd}, or the causally consistent store
of~\cite{gotsmanpopl16}, etc).  Our key technical contribution is thus
a set of proof rules that demonstrate that a program with a given
selection of isolation levels for its transactions preserves its
invariants when executed on a store equipped with appropriate consistency
guarantees.  The paper makes the following contributions:
\vspace*{-6pt}
\begin{enumerate}
  \item We develop a semantics for a core language equipped with
    weakly-isolated transactions, demonstrating that a general
    parametric isolation semantics can be expressed as
    well-formedness constraints on a program's execution.
  \item We present a compositional proof system for this language
    capable of relating high-level application invariants to the structure
    of traces induced by the operational semantics, in which transactions
    are associated with specific weak isolation levels.
  \item We define a \emph{maximum visibility principle} to reconcile
    store consistency guarantees with transaction isolation
    requirements, and use it to instantiate our operational model and
    the proof system for various kinds of (weakly-consistent) data
    stores.
  \item Case studies modeled after real-world scenarios demonstrate
    the applicability and utility of our proof methodology.
\end{enumerate}
\noindent Our results provide the first (to the best of our knowledge)
mechanism that precisely and uniformly relates high-level program
invariants to low-level weak isolation and weak consistency
guarantees, thereby allowing weakly-isolated transactions to enjoy the
same rigorous reasoning capabilities as their strongly-isolated
(serializable) counterparts.
